### 2021-03-29 slides.tex
finish applications paper2 slide

\documentclass[10pt]{beamer}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{bibentry}
\usepackage{biblatex}
\usepackage{adjustbox}
\usepackage[font=scriptsize]{caption}
\usepackage{subcaption}
\captionsetup[figure]{labelsep=period}
\captionsetup[subfigure]{labelformat=simple} % default is 'parens'
\renewcommand\thesubfigure{\thefigure.\alph{subfigure}.}

\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{caption}[numbered]
\bibliography{Bibliography}

\graphicspath{{images/}}
\setbeamertemplate{footline}[frame number]
\setbeamertemplate{page number in head/foot}[appendixframenumber]
\usetheme{Darmstadt}

\title[short]{INTELLIGENT SYSTEMS}
\author[Flavio Forenza]{Flavio Forenza \\ \tiny flavio.forenza@studenti.unimi.it\\[5mm] \includegraphics[scale = 0.06]{logoUnimi2.png}}
\institute{Department of Computer Science,\\ University of Milan, Italy}
\date{\tiny \today}

\begin{document}

\begin{frame}
    \maketitle
\end{frame}

\logo{\includegraphics[width=0.1\linewidth]{logoUnimi2.png}}
\section{Paper 1}
\subsection{\emph{"A Unified Framework for Salient StructureDetection by Contour-Guided Visual Search"}}
\begin{frame}{INTRODUCTION}
    The purpose of the paper is to introduce a new method, called CGVS in the 
    state of the art, which is able to identify regions and salient objects 
    within a scene simultaneously. The proposed system attempts to 
    bridge the gap between the two highly related tasks of \emph{"human 
    fixation"} prediction and \emph{"salient object detection"}, with a general 
    framework.
\end{frame}

\begin{frame}{RELATED WORK}
    To obtain the final result, two pathways are crossed (Fig.\ref{fid: flowchart}):
    \begin{enumerate}
        \item Selective
        \item Non-Selective
    \end{enumerate}
    \begin{figure}[htbp]
        \centering
        \includegraphics[width = 0.8\linewidth]{images/paper1/selective and non-selective pathways.png}
        \centering
        \caption{The flowchart of the porposed system.}
        \label{fid: flowchart}
    \end{figure}
\end{frame}

\begin{frame}[t]{CONTOUR-GUIDED VISUAL SEARCH MODEL pt1}
    In \emph{Non-Selective} pathway are computed:
    \begin{enumerate}
        \item \emph{Boundary} \footfullcite{0747815570}
        \item \emph{Contour-Based Spatial Prior (CBSP)} $ \rightarrow S_w = S_e + S_c $
    \end{enumerate}
    \begin{figure}[htbp]
        \centering
        \includegraphics[width = 0.5\linewidth]{images/paper1/CBSP.png}
        \centering
        \caption{CBSP reconstruction.}\vspace{0mm}
        \label{fid: CBSP}
    \end{figure}
    In \emph{Selective} pathway are computed the basic low-level features:
    \begin{enumerate}
        \item \emph{Luminance} $ \rightarrow f_{lum} = (r+g+b)/3 $
        \item \emph{Color-Opponent} $ \rightarrow f_{rg} = r-g , f_{by} = b-(r+g)/2 $
        \item \emph{Texture Channel} ($ f_{ed} $)
    \end{enumerate}
\end{frame}

\begin{frame}[t]{CONTOUR-GUIDED VISUAL SEARCH MODEL pt2}
    \begin{block}{Bayesian inference}
        All the previous informations are used to compute the Bayesian inference:
        $$ p(s|x) = \frac{p(s)p(x|s)}{p(s)p(x|s)+p(b)p(x|b) } $$
    \end{block}
    Where:
    \begin{enumerate}
        \item $ p(x|s) \rightarrow $ likelihood of a pixel at \emph{x} belonging to a salient structure \emph{s}
        \item $ p(x|b) \rightarrow $ likelihood of a pixel at \emph{x} belonging to the background \emph{b}
        \item $ p(s) \mbox{ and } p(b)=1-p(s) \rightarrow $ prior probabilities of a pixel at x belonging to a salient structure and the background, respectively.
    \end{enumerate}
\end{frame}

\begin{frame}{CONTOUR-GUIDED VISUAL SEARCH MODEL pt3}
    To compute the prior probabilities, we need:
    \begin{enumerate}
        \item \emph{Predict the Size of Potential Structure}: using a optimal threshold ($ T_{opt} $) 
        useful for separating the background from the foreground (salient structures).
        \item \emph{Evaluate the Importance of Each Feature}: weighing every salient structure.
        \item \emph{Calculate the Observation Likelihood:} both $ p(x|s) \mbox{ and } p(x|b) $
        \item \emph{Enhance the Salient Structure by Iterating}: we re-initialize the prior 
        function with the smoothed version of $ p(s|x) $ (by median filtering 
        with a size of 21 × 21 pixels)
    \end{enumerate}
\end{frame}

\begin{frame}[t]{EXPERIMENTS}
    The results obtained on different datasets \footfullcite{PASCAL-S, ECSSD, ASD}, were compared with those 
    btained by the systems already existing at the state of the art.
    \begin{figure}[htbp]
        \centering
        \includegraphics[width = 0.8 \linewidth]{/paper1/metrics.png}
        \centering
        \caption{Metrics comparison.}
        \label{fig: metrics}
    \end{figure}
\end{frame}

\begin{frame}{DISCUSSION}
    It is necessary to say that the results obtained by the proposed method 
    are obtained using the center bias, which if removed places the system 
    performance under those of systems such as \emph{RF} \footfullcite{0747815518} and \emph{HS}\footfullcite{0747815508} (which do not 
    use the central bias). This method offers a dual use. In 
    a cluttered scene with no dominant objects, the system will function as a 
    fixation prediction model. Finally, on the other hand, with simple scenes, 
    the system detects existing relevant objects.
    \begin{figure}[htbp]
        \centering
        \includegraphics[width = 0.7 \linewidth]{paper1/SystemsComp.png}
        \centering
        \label{fig: metrics}
    \end{figure}
\end{frame}

\section{Paper 2}
\subsection{\emph{"Linear Spectral Clustering Superpixel"}}
\begin{frame}{INTRODUCTION}
    The introduced technique is called SUPERPIXEL. Widely used in image 
    processing for particular tasks such as image segmentation, image analysis, 
    image classification, target tracking, 3D reconstruction, surface retrieval and 
    object proposal. The purpose of the elaborate story is to reduce the 
    computational complexity through a superpixel system called LSC (Linear Spectral
    Clustering).
\end{frame}

\begin{frame}{LSC SUPERPIXEL pt1}
    Goal: optimization (max/min) of two objective functions to create clusters of pixels called superpixels:
    \begin{block}{Objective Function 1: Weighted K-Means}
        $$ F_{km} = \sum_{k=1}^K\sum_{p\in\pi_k}\omega(p)= || \phi(p)-m_k ||^2 $$
    \end{block}
    \begin{block}{Objective Function 2: Normalized cuts}
        $$ F_{N_{cuts}} = \frac{1}{K}\sum_{k=1}^K\frac{\sum_{p\in\pi_k}\sum_{q\in\pi_k}W(p,q)}{\sum_{p\in\pi_k}\sum_{q\in{V}}W(p,q)} $$
    \end{block}
    Problem: extremely high computational!
\end{frame}

\begin{frame}{LSC SUPERPIXEL pt2}
    If corollary 1 is satisfied, then  \emph{weghted K-means clustering} can be used 
    for segmentation and for creation of superpixel regions, instead of the 
    eigen-vector based method.
    \begin{block}{\bfseries{Corollary 1}}
        Optimizations of the objective functions of the weighted K-means
        and the normalized cuts are mathematically equivalent if (1) and (2) 
        hold simultaneously. The symbol $ \cdot $ stands for inner product.
    \end{block}

    \begin{block}{Equation (1)}
           $$ \omega(p)\phi(p) \cdot \omega(q)\phi(q) = W(p,q), \forall p,q \in V $$
    \end{block}

    \begin{block}{Equation (2)}
        $$ \omega(p) = \sum_{q \in V} W(p,q), \forall p \in V $$
    \end{block}
\end{frame}

\begin{frame}{LSC Algorithm pt1}
    Problem? $ \rightarrow $ Find the correct positive similarity function $ W (p,q) $, 
    between two data points \emph{p} and \emph{q} to solve equation 1.
    \begin{block}{Similarity Function}
        \begin{equation}\small
            \begin{split}
                W(p,q) = C_s^2(\cos \frac{\pi}{2}(x_p-x_q)+\cos\frac{\pi}{2}(y_p-y_q)) \\
                + C_c^2(\cos \frac{\pi}{2}(l_p-l_q)+\cos\frac{\pi}{2}(\alpha_p-\alpha_q) \\
                + \cos\frac{\pi}{2}(\beta_p-\beta_q)x2.55^2)
            \end{split}
        \end{equation}
    \end{block}

    \begin{block}{Mapping Function}
        \begin{equation}\small
            \begin{split}
                \phi(p) = \frac{1}{\omega(p)}[C_c\cos\frac{\pi}{2}l_p, C_c\sin\frac{\pi}{2}l_p, 2.55C_c\cos\frac{\pi}{2}\alpha_p \\
                x 2.55C_c\sin\frac{\pi}{2}\alpha_p, 2.55C_c\cos\frac{\pi}{2}\beta_p, 2.55C_c\sin\frac{\pi}{2}\alpha_p, \\
                x C_s\cos\frac{\pi}{2}x_p, C_s\sin\frac{\pi}{2}x_p, C_s\cos\frac{\pi}{2}y_p, C_s\sin\frac{\pi}{2}x_p]
            \end{split}
        \end{equation} 
    \end{block}

\end{frame}

\begin{frame}{LSC Algorithm pt2}
    Assuming that equation 1 exists, LSC method takes two parameters as 
    input; The image to be segmented and the preferred K number of 
    superpixels that the algorithm will have to produce. The number K 
    corresponds to the number of centroids and each of these will be useful 
    for grouping the neighbors pixels, in a range of $ \tau v_x \ x \ \tau v_y $, with $ \tau>0.5 $, 
    using a distance comparison.
    \begin{figure}[h!]
        \centering
        \includegraphics[width = 0.7 \linewidth]{paper2/slide1.png}
        \centering
        \caption{Sample images segmented with K = 1000/500/200 superpixels using LSC.}
        \label{fig: metrics}
    \end{figure}
\end{frame}

\begin{frame}{COMPARATIVE EXPERIMENTS}
    \begin{table}[htbp!]
        \centering
        \begin{adjustbox}{max width=\textwidth}
        \begin{tabular}{*{9}{|c}|}%%{|c|c|c|c|c|c|c|c|c|}
            \hline
            & EneOpt0 & SEEDS\footfullcite{0781426509} & ERS\footfullcite{0781426508} & Lattices & NCuts & SLIC\footfullcite{0781426514} & Turbo & LSC \\
            \hline
            \bfseries{ADERENCE TO BOUNDARY} & & & & & & & & \\
            \emph{Under segmentation error} & 0.230 & 0.197 & 0.198 & 0.303 & 0.220 & 0.213 & 0.277 & \bfseries{0.190}\\
            \emph{Boundary recall} & 0.765 & 0.918 & 0.920 & 0.811 & 0.789 & 0.837 & 0.739 & \bfseries{0.926}\\
            \emph{Achievable segmentation accuracy} & 0.950 & 0.960 & 0.959 & 0.933 & 0.956 & 0.956 & 0.943 & \bfseries{0.962}\\
            \hline
            \bfseries{SEGMENTATION SPEED} & & & & & & & & \\
            \emph{Computational complexity} & $ O(N^3/K^2) $ & $ O(N) $ & $ O(N^2 \lg{N}) $ & $ O(N^{\frac{3}{2}} \lg{N}) $ & $ O(N^{\frac{3}{2}}) $ & $ O(N) $ & $ O(N) $ & $ O(N) $\\
            \emph{Average time per image} & 3.35s & \bfseries{0.0935}s & 0.969s & 0.284s & 93.4s & 0.125s & 6.61s & 0.334s\\
            \hline
        \end{tabular}
        \end{adjustbox}
        \caption{Performance metrics superpixel segmentation algorithms at K=400}
        \label{table superpixels}
    \end{table}
    Best performance: low "Under segmentation Error" (\emph{UE}), high 
    "Boundary Recall" (\emph{BR}) and high "Achievable segmentation accuracy 
    (\emph{ASA}).
\end{frame}

\begin{frame}{APPLICATIONS: Class segmentation}
    \begin{figure}[htbp]
        \centering
        \includegraphics[width = 1 \linewidth]{images/paper2/superpixelAlgo.png}
        \centering
        \caption{Segmentation using different superpixels algorithms. (a) Original Image. (b) QS. (c) ERS. (d) SLIC. (e) LSC. (f)Ground Truth.}
        \label{fig: superpixelSegmentation}
    \end{figure}

    \begin{table}[h!]
        \centering
        \begin{adjustbox}{max width=4cm}
        \begin{tabular}{*{5}{|c}|}%%{|c|c|c|c|c|}
            \hline
            & QS & ERS & SLIC & LSC\\
            \hline
            bike & 72.2 & 74.2 & 76.3 & \bfseries{76.9}\\
            cars & 72.2 & 74.7 & 72.5 & \bfseries{76.8}\\
            person & 66.3 & 66.5 & 66.7 & \bfseries{67.0}\\
            \hline
        \end{tabular}
        \end{adjustbox}
        \caption{Accuracy using different superpixels algorithms.}
        \label{table accuracy}
    \end{table}
\end{frame}

\begin{frame}{APPLICATIONS: Weakly Supervised Semantic Segmentation}
    \begin{minipage}{\linewidth}
        \centering
        \begin{minipage}{0.45\linewidth}
            \begin{figure}[H]
                \includegraphics[width = 0.7 \linewidth]{images/paper2/semanticSegmentation.png}
                \caption{Weakly supervised semantic segmentation. (a) A training image with bounding boxes. (b) Output of soft-max layer of $ FCN_c $. (c) Coarse semantic segmentation result. (d) Superpixels with higher probability of foreground. (e) Fore-/Background segmentation result after iterative optimization. (f) Refined semantic segmentation result.}
            \end{figure}
        \end{minipage}
        \hspace{0.05\linewidth}
        \begin{minipage}{0.45\linewidth}
            \begin{figure}[htbp]
                \centering
                \includegraphics[width = 1 \linewidth]{images/paper2/segmentationAlgo.png}
                \centering
                \caption{Semantic segmentation. (a) Input image. (b) Strong. (c) Bbox-seg. (d) $ Joint_{sp} $ (LSC) }
            \end{figure}
            \begin{table}[H]
                \begin{adjustbox}{max width=4cm}
                    \begin{tabular}{*{3}{|c}|}%%{|c|c|c|}
                        \hline
                        Strong\footnotemark[1]& Bbox-seg\footnotemark[1] & $ Joint_{sp} $ \\
                        \hline
                        62.5 & 60.6 & \bfseries{64.0} \\
                        \hline
                    \end{tabular}
                \end{adjustbox}
                \caption{Semantic segmentation accuracy in terms of Mean IOU (\%)}
            \end{table}
        \end{minipage}
    \end{minipage}
    \footnotetext[1]{\tiny G. Papandreou, L. Chen, K. Murphy, and A. Yuille, “Weakly-and semi-
    supervised learning of a deep convolutional network for semantic image
    segmentation,” in Proc. ICCV, pp. 1742–1750, Dec. 2015}

\end{frame}

\end{document}