### 2021-03-29 paper_3.tex
start paper3

\section{An End-to-End Compression Framework Based
on Convolutional Neural Networks}

\begin{flushleft}
    \author{
    Feng Jiang, 
    Wen Tao, 
    Shaohui Liu, 
    Jie Ren, 
    Xun Guo, 
    Debin Zhao, 
    \emph{Member, IEEE}
    }
\end{flushleft}

\begin{center}
    \emph{IEEE TRANSACTIONS ON CIRCUIT AND SYSTEMS FOR VIDEO THECNOLOGY, VOL. 28, NO. 11, OCTOBER 2018}
\end{center}

\subsection{INTRODUCTION}