### 2021-04-01 paper_4.tex
start introduction paper4

\section{A Data Set for Camera-Independent Color Constancy}

\begin{flushleft}
    \author{
    Ça$ \breve{g} $lar Aytekin, 
    Jarno Nikkanen, 
    Moncef Gabbouj
    \emph{Fellow, IEEE}
    }
\end{flushleft}

\begin{center}
    \emph{IEEE TRANSACTIONS ON IMAGE PROCESSING, VOL. 27, NO. 2, FEBRUARY 2018}
\end{center}

\subsection{INTRODUCTION}