### 2021-04-10 paper_6.tex
start paper6

\section{Coarse-to-Fine Semantic Segmentation From Image-Level Labels}

\begin{flushleft}
    \author{
    Longlong Jing,
    Yucheng Chen,
    Yingli Tan,
    \emph{Fellow, IEEE}
    }
\end{flushleft}

\begin{center}
    \emph{IEEE TRANSACTIONS ON IMAGE PROCESSING, VOL.29, 2020}
\end{center}

\subsection{INTRODUCTION}