### 2021-04-10 paper_6.tex
finish 6.2

\section{Coarse-to-Fine Semantic Segmentation From Image-Level Labels}

\begin{flushleft}
    \author{
    Longlong Jing,
    Yucheng Chen,
    Yingli Tan,
    \emph{Fellow, IEEE}
    }
\end{flushleft}

\begin{center}
    \emph{IEEE TRANSACTIONS ON IMAGE PROCESSING, VOL.29, 2020}
\end{center}

\subsection{INTRODUCTION}
To obtain a semantic segmentation it is necessary to have an annotated 
training set. Semantic segmentation generally requires pixel-wise semantic 
annotation techniques which are time-consuming. To solve the annotation 
problem, weakly supervised and semi-supervised semantic segmentation 
methods have been proposed. Methods such as object-level labels and image-level 
labels are used to extract annotations. These annotations are useful for 
producing datasets useful for training a weakly supervised model, such as 
the one proposed. By training a weakly supervised semantic segmentation 
model, with the labels produced, the performances match those of the supervised 
methods. Unfortunately, methods such as object-level labels produce 
bounding boxes that are still not precise, so the use of the image-level labels 
method prevails. The dataset used is that of ImageNet where a category is 
associated with each image. The proposed model is able to generate as many 
segmentation masks as there are different objects belonging to different 
categories in each image. Network training takes place using information such as 
masks, images and labels of each category. An example of the coarse-to-fine 
mask produced by the model is the one shown in figure \ref{fig:semanticWork}.
\begin{figure}[h!]
    \centering
    \includegraphics[width = 1 \linewidth]{images/paper6/work.png}
    \centering
    \caption{Example of semantic segmentation generated by the proposed method.}
    \label{fig:semanticWork}
\end{figure}

\subsection{RELATED WORK}
\subsubsection{Semantic Segmentation}
There are three different types of semantic segmentation methods that differ 
based on the level of annotation used:
\begin{enumerate}
    \item \emph{Fully Supervised Pixel-Wise Annotation-Based Methods}: trained with 
    pixel-wise labels annotated by people:
    \item \emph{Weakly Supervised Object-level Annotation-Based-Methods}: trained with 
    methods such as object-level annotations based on the use of bounding 
    boxes;
    \item \emph{Weakly Supervised Image-Level Annotation-Based Methods}: trained 
    with image category labels.
\end{enumerate}
The methods in the first point, if trained with precise labels, are able to 
obtain the best performances.

\subsubsection{Foreground Segmentation}
The labels assigned to pixels can be useful to separate a foreground object from a background object. Also for this task there are three strategies to perform a segmentation of the objects in the foreground:
\begin{enumerate}
    \item \emph{Joint segmentation-based Methods}: used prior knowledge as supervision.
    \item \emph{Saliency prediction-based Methods}: identify regions present in the human visual scene.
    \item \emph{Object proposal-based Methods}: locate all objects in images.
\end{enumerate}
The proposed framework can also be used for this purpose.