### 2021-04-11 slides/slides_paper6.tex

\section{Paper 6}
\subsection{\emph{"Coarse-to-Fine Semantic Segmentation From Image-Level Labels"}}

\begin{frame}{INTRODUCTION}
    ciao mondo
\end{frame}