### 2021-04-12 papers/paper_7.tex
start paper7

\section{Variational-Based Mixed Noise Removal With CNN Deep Learning Regularization}

\begin{flushleft}
    \author{
    Faqiang Wang,
    Haiyang Huang,
    Jun Liu
    }
\end{flushleft}

\begin{center}
    \emph{IEEE TRANSACTIONS ON IMAGE PROCESSING, VOL.29, 2020}
\end{center}

\subsection{INTRODUCTION}