### 2021-04-16 papers/paper_8.tex
finish 8.3.4

\section{Visual Object Tracking via Multi-Stream Deep Similarity Learning Networks}

\begin{center}
    \author{
    Kunpeng Li,
    \emph{Student Member, IEEE},
    Yu Kong,
    \emph{Member, IEEE},
    Yunf Fu,
    \emph{Fellow, IEEE}
    }
\end{center}

\begin{center}
    \emph{IEEE TRANSACTIONS ON IMAGE PROCESSING, VOL.29, 2020}
\end{center}

\subsection{INTRODUCTION}
The goal of visual tracking systems is to be able to obtain the position of 
the tracked target even in subsequent frames. The problems to be solved are 
those of occlusion, background clutter, illumination variations, deformation 
etc. The existing state-of-the-art models carry out training online exclusively. 
Some methods, based on pre-trained convolutional networks (CNN), track 
the object based on the background. In these networks, stochastic gradient 
descent is applied in order to update the entire network. However, these 
models appear to be too slow and do not work well in real-time. The following 
paper uses a model that compares similarities and is trained offline in order to 
predict the patch present in the next frame. The method, thanks to the use of 
the relative distance, is robust in the presence of phenomena that introduce 
the so called distractors. The purpose is to be able to compare the patch that 
identifies the target template (the object) with all the possible positions 
of the same region belonging to the next frame. In addition to tracking the 
object, the proposed model is a framework for EMDSLT tracking like the 
one shown in figure \ref{fig:EMDSLT}. Within it, two procedures are considered important: 
updating the model and self-recovery from failures. The network is composed 
of two types of structures, one that is responsible for dast-speed verification 
and production of the tracking results, while the second is responsible for re-verification 
and re-detection of the target object on the patches previously 
generated.
\begin{figure}[h!]
    \centering
    \includegraphics[width = \linewidth]{images/paper8/EMDSLT.png}
    \centering
    \caption{The framework of EMDSLT proposed method.}
    \label{fig:EMDSLT}
\end{figure}.

\subsection{RELATED WORK}
Methods such as generative trackers have been developed for finding positions 
in which to find the best candidate during the tracking process. Good 
results have been achieved by compound trackers \cite{0893551105} \cite{0893551118}. Networks such as 
RNNs \cite{0893551121} can be used to predict possible target positions in each frame. 
However, these types of networks have not yet achieved satisfactory results. 
Some methods instead train a CNN offline while the test is done online. Unfortunately, 
these models are not fast. Other models instead if they fail a first 
time, then they will always fail until the target returns to the search region. 
Methods such as \cite{0893551129} \cite{0893551134} search for the target template in subsequent frames 
in the same position, obtaining better performance than methods based on 
similarity but nevertheless are not robust to distracting elements. The proposed 
method instead takes into account both the background patches and 
the target template patches in order to make comparisons on similarities.

\subsection{METHODS}
\subsubsection{Multi Stream Deep Similarity Learning Networks}
A mulit-stream similarity learning network is used in this paper. In the 
proposed framework there are various data streams (Fig. \ref{fig:EMDSLT}):
\begin{enumerate}
    \item Template Stream: the target template defined in the beginning frame;
    \item Searching Stream: the search region;
    \item Background Stream: background patches that sample around the target 
    in the beginning frame.
\end{enumerate}

\subsubsection{Template-Searching-Background Loss}
Several positive $X^+_i$, negative $X^-_i$ and background $B_{i-1,j}$ patches can be 
generated in an image. Patches are considered positive if they have a small 
cosine distance with the target template $T$, on the contrary, patches will be 
considered negative if they are closer to the background patches within the 
search region $S_i$. The loss function to be minimized is the following:
\begin{equation}
    L(T,\{X_{i,k}\}^N_{k=1}, \{B_{i-1}\}^M_{j=1}) = \sum_{k=1}^NL(T,X_{i,k}, \{B_{i-1,j}\}^M_{j=1})
\end{equation}
Where $T$ is the target template, $X_{i,k}$ is the $k-th$ candidate patch $X$ of frame 
$i$, $M$ is the number of background patches and $N$ is the number of candidate 
patches $X_{i,k}$.

\subsubsection{Tracking Method}
Background patches are obtained by sampling around the previous positions 
of frame $i-1$. The background patches that have a high correlation with 
the target template will be definitively chosen. Find these patches first, the 
selected $X_i$ pateches will be the ones that will match the target template and 
discriminate against the background patches. Ultimately, the choice of 
patch $X_i$ comes down to an optimization problem.
\begin{equation}
    X_i = \argmin_{X_{i,k} \in S_i}d(T,X_{i,k}, \{B_{i-1,j}\}^M_{j=1})
\end{equation}
where $d(\cdot)$ is a distance function.

\subsubsection{Tracking With Updating}
As already mentioned, the two important aspects of the framework are to update the model and allow self-recovery when it is unable to track the target. In order to guarantee these two operations, a confidence assessment of the monitoring results must be carried out. The set of patches Xi, each of which belongs to each frame ($K = k_1$), is passed to the part of the network that performs the re-verification in order to return the "\emph{Trust Score s}" obtained based on the similarity with the target template . This score indicates the confidence level of the monitoring and can then be used to identify any errors that can be corrected with an update of the taget template. When the s score is less than a $\sigma$ threshold. then a failure has occurred. If a failure occurs, then the number of frames ($K = k_2$) would be increased by 1 and then re-detected, on the patches, which would return a new patch $X'$ with a new score $s'$. If $s'> \sigma$, then this will be the new tracking result ending the recovery process. When this occurs, the target template, calculated with the new formula (\ref{newX}) will be updated and then replaced with the last detected. In \ref{newX}, $m$ represents the matching function, $m(a,b)=f(a)^Tf(b)$, used to obtain the Trust Score. 
\begin{equation}\label{newX}
    X'_i = \argmax_{X_{i,k} \in S_i}m(T,X_{i,k})
\end{equation}