### 2021-04-17 papers/paper_9.tex
start paper9

\section{Similar Face Recognition Using the IE-CNN Model}

\begin{center}
    \author{
    An-Ping Song,
    Qian Hu,
    Xue-Hai Ding,
    Xin-Yi Di,
    Zi-Heng Song
    }
\end{center}

\begin{center}
    \emph{DIGITAL OBJECT IDENTIFIER}
\end{center}

\subsection{INTRODUCTION}